---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(dplyr)
library(stringr)
library(ggsci)
library(RColorBrewer)
library(GGally)
```

```{r}
getwd()
```


## Reading the original file (no curation)

```{r}
#file <- read.csv("../../01_preprocessing/data/ECO-data-set-dl-training-test-clean-unique-sentences-rCFMC_v3.tsv", #header= TRUE, sep = "\t")

```

## Reading eco file  
```{r}
file_eco <- read.csv("../../01_preprocessing/results/ECO/curated_master_dataset/ECO_dataset_master_curated_dataset_v2.tsv", header= TRUE, sep = "\t")
```

## Reading salmonella file 

```{r}
file_salmo <- read.csv("../../../Data-sets/STM-data-set-dl-sentence-effect.tsv", header = TRUE, sep = "\t")
```


## Counting eco unique tfs and regulated genes

```{r}
count_of_eco_regulators<- length(unique(file_eco$REGULATOR))
print(count_of_eco_regulators)
count_of_eco_regulated<- length(unique(file_eco$REGULATED))
print(count_of_eco_regulated)
```

## Counting salmonela unique tfs and regulated genes

```{r}
count_of_salmo_regulators<- length(unique(file_salmo$REGULATOR))
print(count_of_salmo_regulators)
count_of_salmo_regulated<- length(unique(file_salmo$REGULATED))
print(count_of_salmo_regulated)
```

# Bar plot visualization

## Generating the dataframe needed for barplot visualization 

```{r}
eco_norm_effect_df<- data.frame(table(file_eco$NORMALIZED_EFFECT))
colnames(eco_norm_effect_df)<- c("effect","examples")
```

```{r}
salmo_norm_effect_df <- data.frame(table(file_salmo$EfectoNormalizado))
colnames(salmo_norm_effect_df)<- c("effect","examples")
```

# Setting a color blind friendly pallette
```{r}
effect_unique<- unique(eco_norm_effect_df$effect)
```

```{r}
# Function to generate color mapping dataframe
get_color_mapping <- function(categories) {
  colors <- brewer.pal(length(categories), "Dark2")
  color_mapping_df <- data.frame(
    effect = categories,
    color = colors
  )
  return(color_mapping_df)
}

# Generate color mapping based on the Dark2 palette
color_mapping <- get_color_mapping(effect_unique)

# Merge color mapping with dataframes
eco_df1 <- merge(eco_norm_effect_df, color_mapping, by = "effect")
salmo_df1 <- merge(salmo_norm_effect_df, color_mapping, by = "effect")

```

## Sorting the categories 

```{r}
eco_df1_sorted <- eco_df1[order(-eco_df1$examples), ]
salmo_df1_sorted <- salmo_df1[order(-salmo_df1$examples), ]
```



## Creating the barplot for eco

```{r}

ggplot(eco_df1_sorted, aes(x=reorder(effect, -examples), y=examples,fill=color)) +
  geom_bar(stat="identity") +
  #ggtitle("Relation Distribution of Entities") +
  xlab("Category") +
  ylab("Number of Examples") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  scale_fill_identity() +
  #scale_fill_brewer(palette= "Dark2") +
  geom_text(aes(label=examples), vjust=0)

```

```{r}
ggsave("../results/computers_in_biology_paper/eco/eco_relation_distribution_of_entities_for_paper_v4.jpeg",device ="jpeg", dpi =300)
```
## Creating the barplot for salmonella

```{r}
ggplot(salmo_df1_sorted, aes(x=reorder(effect, -examples), y=examples,fill=color)) +
  geom_bar(stat="identity") +
  #ggtitle("Relation Distribution of Entities") +
  xlab("Category") +
  ylab("Number of Examples") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  scale_fill_identity() +
  # Dark2 is color blind friendly palette
  geom_text(aes(label=examples), vjust=0)

```

```{r}
ggsave("../results/computers_in_biology_paper/salmo_264/salmo_relation_distribution_of_entities_for_paper_v4.jpeg",device ="jpeg", dpi =300)
```


# Sentence length distribution

## Sentence length stats 

```{r}
file_eco$sentence_length<- str_length(file_eco$SENTENCE)
summary(file_eco$sentence_length)
```

```{r}
file_salmo$sentence_length<- str_length(file_salmo$SENTENCE)
summary(file_salmo$sentence_length)
```


## Generating the dataframe needed for histogram visualization of eco 

```{r}
# Color-blind friendly 
color_palette <- brewer.pal(8, "Set1")

```

```{r}
library(RColorBrewer)
display.brewer.pal(n = 8, name = "Set1")
```


```{r}
ggplot(file_eco, aes(x=sentence_length)) + 
  geom_histogram(color=color_palette[7], fill= alpha(color_palette[2], 0.7),aes(y=(..count..)/sum(..count..)),binwidth=50) +
  xlab("Sentence Length") +
  ylab("Percentage of Dataset") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent)
```

```{r}
ggsave("../results/computers_in_biology_paper/eco/eco_sentence_length_distribution_outliers_for_paper_v3.jpg", device = "jpeg", dpi=300)
```

## Generating the dataframe needed for histogram visualiztion of salmonella


```{r}
ggplot(file_salmo, aes(x=sentence_length)) + 
  geom_histogram(color=color_palette[7], fill= alpha(color_palette[2], 0.7),aes(y=(..count..)/sum(..count..)),binwidth=50) +
  xlab("Sentence Length") +
  ylab("Percentage of Dataset") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent)
```

```{r}
ggsave("../results/computers_in_biology_paper/salmo_264/salmo_sentence_length_distribution_outliers_for_paper_v3.jpg", device = "jpeg", dpi=300)
```




## Zooming into the large distribution by ignoring greater than 1000 elements 

```{r}
salmo_less_1500<- file_salmo %>% filter(sentence_length<1500)

```


```{r}
summary(salmo_less_1500$sentence_length)
```

```{r}
ggplot(salmo_less_1500, aes(x=sentence_length)) + 
  geom_histogram(color=color_palette[7], fill= alpha(color_palette[2], 0.7),aes(y=(..count..)/sum(..count..)),binwidth=50) +
  xlab("Sentence Length") +
  ylab("Percentage of Dataset") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent)
```


```{r}
ggsave("../results/computers_in_biology_paper/salmo_264/salmo_sentence_length_distribution_no_outliers_less_than_1500_for_paper_v3.jpg", device = "jpeg", dpi=300)
```


## Counting different PMID in the sentence dataset for ecoli

```{r}
eco_different_pmids_count<- length(unique(file_eco$PMID))
print(eco_different_pmids_count)
```

## Counting different PMID in the sentence dataset for salmonella

```{r}
salmo_different_pmids_count<- length(unique(file_salmo$PMID))
print(salmo_different_pmids_count)
```


## List of different ecoli PMIDs 
```{r}
eco_list_of_unique_pmids<- unique(file_eco$PMID)
print(eco_list_of_unique_pmids)

```

## List of different salmonella PMIDs 
```{r}
salmo_list_of_unique_pmids<- unique(file_salmo$PMID)
print(salmo_list_of_unique_pmids)

```

## Different pmids viz 

### For ecoli 
```{r}
eco_pmid_df<- data.frame(table(file_eco$PMID))
colnames(eco_pmid_df) <- c("pmid","examples")
eco_pmid_df<- eco_pmid_df[order(eco_pmid_df$examples,decreasing = TRUE),]
```

### For salmonella

```{r}
salmo_pmid_df<- data.frame(table(file_salmo$PMID))
colnames(salmo_pmid_df) <- c("pmid","examples")
salmo_pmid_df<- salmo_pmid_df[order(salmo_pmid_df$examples,decreasing = TRUE),]
```



## Palette creation for 119 colours of ecoli and 264 of salmo, always color-blind

```{r}
brewer.pal.info
```


```{r}
colors_119<- colorRampPalette(c("red","green", "blue"),space="Lab")(eco_different_pmids_count) 
```

```{r}
eco_colour_count<- eco_different_pmids_count 
salmo_colour_count <- salmo_different_pmids_count

#get_palette_accent<- colorRampPalette(brewer.pal(12,"Accent"))(colour_count)
#get_palette_spectral<- colorRampPalette(brewer.pal(11,"Spectral"))(colour_count)
#get_palette_blues <- colorRampPalette(brewer.pal(9,"Blues"))(colour_count)
#get_palette_dark2<- colorRampPalette(brewer.pal(8,"Dark2"))(colour_count)
#get_palette_set2<- colorRampPalette(brewer.pal(8,"Set2"))(colour_count)
get_eco_palette_paired<- colorRampPalette(brewer.pal(12,"Paired"))(eco_colour_count)
get_salmo_palette_paired<-colorRampPalette(brewer.pal(12,"Paired"))(salmo_colour_count)

```


## Creating the barplot for ecoli 

```{r}
ggplot(eco_pmid_df, aes(x=reorder(pmid,examples), y=examples,fill=pmid)) +
  geom_bar(stat="identity") +
  theme_gray() +
  #ggtitle("PMID Distribution of Sentences") +
  xlab("PMID") +
  ylab("Percentage of Dataset") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5), axis.text.y = element_text(size=4)) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1, scale = .1)) +
  scale_fill_manual(values= get_eco_palette_paired)
```

```{r}
ggsave("../results/computers_in_biology_paper/eco/eco_pmid_distribution_of_sentences_for_paper_v1.jpg",device="jpeg",dpi=300)
```

## Creating the barplot for salmonella 

```{r}
ggplot(salmo_pmid_df, aes(x=reorder(pmid,examples), y=examples,fill=pmid)) +
  geom_bar(stat="identity") +
  theme_gray() +
  #ggtitle("PMID Distribution of Sentences") +
  xlab("PMID") +
  ylab("Percentage of Dataset") + 
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5), axis.text.y = element_text(size=3)) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1, scale = .1)) +
  scale_fill_manual(values= get_salmo_palette_paired)
```



```{r}
ggsave("../results/computers_in_biology_paper/salmo_264/salmo_pmid_distribution_of_sentences_for_paper_v1.jpg",device="jpeg",dpi=300)
```

# Parallel coordinate plot for biolinkbert sweep 


```{r}
file_biolinkbert <- read.csv("../../03_visualization/data/biolinkbert_sweep_for_parallel_coords_viz.csv", header = TRUE)

```

```{r}
file_biolinkbert$Name <- as.factor(file_biolinkbert$Name)
str(file_biolinkbert)
```


```{r}
ggparcoord(file_biolinkbert,
    columns = 1:6,groupColumn = "validation_f1_epoch",showPoints = TRUE, splineFactor = TRUE
    ) 
```


```{r}
parcoords(file_biolinkbert)
```




